Run the `.xsl` style sheet as shown in `make-data.sh` on the `.xml` files. Read the `.csv ` files into R and `save()` as `.rda` files.

```{r}
library(dplyr)
```

To read and save the .rda files in Dropbox, first set your working
directory to the wasp-trips/data folder before running this code:

Calculate time differences to correct the recorded times:

wasp1 should be 7 hours earlier:

```{r}
actual_time_w1 <- as.POSIXct(strptime("01/01/2013 01:00:00", format
  = "%m/%d/%Y %H:%M:%S"))
preset_time_w1 <- as.POSIXct(strptime("01/01/2013 08:00:00", format
  = "%m/%d/%Y %H:%M:%S"))
w1_add_time <- actual_time_w1 - preset_time_w1
```

wasp2 should be quite a bit later:
```{r}
actual_time_w2 <- as.POSIXct(strptime("11/05/2013 14:47:10", format
  = "%m/%d/%Y %H:%M:%S"))
preset_time_w2 <- as.POSIXct(strptime("02/11/2005 05:50:34", format
  = "%m/%d/%Y %H:%M:%S"))
w2_add_time <- actual_time_w2 - preset_time_w2
```

Now process the 2 sets of data:

```{r, eval=FALSE}
load("wasps1.rda")
wasps1_df <- group_by(wasps1, uid)
rm(wasps1)
wasps1_trips <- do(wasps1_df, function(x) get_wasp_trips(x,
    diff_address_arriving = -1, diff_address_leaving = 1,
    add_time = w1_add_time))
rm(wasps1_df)
wasps1_trips <- do.call("rbind", wasps1_trips)
wasps1_trips <- as.data.frame(wasps1_trips)
gc()

load("wasps2.rda")
wasps2_df <- group_by(wasps2, uid)
rm(wasps2)
wasps2_trips <- do(wasps2_df, function(x) get_wasp_trips(x,
    diff_address_arriving = -2, diff_address_leaving = 2,
    add_time = w2_add_time))
rm(wasps2_df)
wasps2_trips <- do.call("rbind", wasps2_trips)
wasps2_trips <- as.data.frame(wasps2_trips)
gc()

wasps1_trips$group <- 1
wasps2_trips$group <- 2
wasp <- rbind(wasps1_trips, wasps2_trips)
rm(wasps1_trips, wasps2_trips)

# convert to POSIXct to use in data frames / ddply / ggplot2
row.names(wasp) <- NULL
wasp$time <- as.POSIXct(wasp$time)

save(wasp, file = "wasp.rda")
```

Calculate hour of day:

```{r}
wasp$time_trunc <- as.POSIXct(trunc(wasp$time, "days"))
wasp$hour_of_day <- as.numeric(difftime(wasp$time, wasp$time_trunc,
    units = "hours"))
```

Now merge in the tag-caste data:

```{r}
load("wasp.rda")
load("cohorts_1.rda")
cohorts_1$caste <- as.character(cohorts_1$caste)
cohorts_1$date_tagged <- as.character(cohorts_1$date_tagged)
cohorts_2 <- read.csv("cohorts_2.csv", stringsAsFactors = FALSE)
cohorts_1$group <- 1
cohorts_2$group <- 2
cohorts <- rbind(cohorts_1, cohorts_2)
cohorts$date_tagged <- strptime(cohorts$date_tagged, format = "%Y-%m-%d")
cohorts$date_tagged <- as.POSIXct(cohorts$date_tagged)

# Create a transformed uid column that matches the tag format:
wasp <- transform(wasp, tag = paste0(substr(uid, 1, 2), substr(uid, 4, 5)))
wasp$uid <- NULL

library(plyr)
wasp_caste <- join(wasp, cohorts, by = c("tag", "group"))
last_trip <- ddply(wasp_caste, c("tag", "caste"), function(x) tail(x, n = 1))
first_trip <- ddply(wasp_caste, c("tag", "caste"), function(x) head(x, n = 1))
```

Calculate age based on the date_tagged cohort data.

```{r}
wasp_caste$age <- wasp_caste$time_trunc - wasp_caste$date_tagged
save(wasp_caste, file = "wasp_caste.rda")
```

And create some summary data:

```{r}
wasp_summary <- ddply(wasp_caste, c("trips", "tag", "caste", "group"), summarize,
  mean_trip = mean(time_diff/60/60, na.rm = TRUE), numb_trips = length(trips),
  mean_trip_sd = sd(time_diff/60/60, na.rm = TRUE),
  median_trip = median(time_diff/60/60, na.rm = TRUE), l_quant =
  quantile(time_diff/60/60, 0.25, na.rm = TRUE), u_quant =
  quantile(time_diff/60/60, 0.75, na.rm = TRUE))

wasp_summary <- transform(wasp_summary, ordered_tag = reorder(tag,
    -median_trip))
save(wasp_summary, file = "wasp_summary.rda")
```
